r5r: rapid realistic routing on multimodal transport networks with R5 in R

### 2020-10-21 r-package/paper2020/paper2020.Rmd
adds travel time matrix example using time window to paper

- `travel_time_matrix()`: used to compute travel time estimates between one or 
multiple origin/destination pairs for a single departure time or for multiple
departure times over a time window. The origin/destination input can be either a 
spatial `sf POINT` object, or a `data.frame` containing the columns `id`, `lon`, 
`lat.` The output is a `data.frame` with the travel time estimate between each 
origin/destination pair. If the user passes a `time_window` parameter to the,
function, r5r will calculate multiple travel time matrices departing each minute
and return additional information on the percentile distribution of travel times
that allows one to grasp service reliability issues.

@@

```{r, message = FALSE, eval = FALSE}
# stable version
install.packages("r5r")

```

@@

#> I think this is perhaps unecessary detail for the paper publication.
Please note that `R` only allocates memory to Java once during a running 
session. If a Java process has been initiated before, either by r5r or any other 
Java-dependent package or function, it might be required to restart the current 
`R` session in order to effectively change the amount of memory allocated.

@@

To illustrate functionality, the package includes a small sample data set for 
the city of Porto Alegre (Brazil). It includes three files:

- An OSM network in `.pbf` format.
- A public transport network in `GTFS.zip` format.
- The spatial coordinates of points covering the area in `.csv` format.

```{r}
data_path <- system.file("extdata", package = "r5r")
list.files(data_path)

points <- fread(system.file("extdata/poa_hexgrid.csv", package = "r5r"))
points <- points[c(sample(1:nrow(points), 100))]
head(points)
```

@@

Calculating a travel time matrix with r5r can be done very efficiently with the
`travel_time_matrix()` function. The function takes travel parameters and
origins/destinations spatial location as inputs and outputs a `data.table` with 
the travel time between all combinations of origins and destinations that can be 
reached with the given parameters:

```{r, message = FALSE}
departure_datetime <- as.POSIXct("13-03-2019 14:00:00",
                                 format = "%d-%m-%Y %H:%M:%S")

# calculate travel time matrix
ttm <- travel_time_matrix(r5r_core,
                          origins = points,
                          destinations = points,
                          mode = c("WALK", "TRANSIT"),
                          departure_datetime,
                          max_walk_dist = 1000,
                          max_trip_duration = 120)

@@

Note that the resulting `ttm` object only holds the travel time of 1043 
origin/destination pairs, even though the `points` data set passed both to 
`origins` and `destinations` argument contains 100 points. This is because out 
of the total 10000 possible trips between the supplied points, only 1043 could 
be completed within the specified restrictions. Removing any walking distance 
limits (by setting it to infinite) and increasing maximum trip duration 
accordingly results in a "complete" output:

```{r, message = FALSE}
# set softer restrictions
max_walk_dist <- Inf
max_trip_duration <- 360

# calculate travel time matrix
ttm <- travel_time_matrix(r5r_core,
                          origins = points,
                          destinations = points,
                          mode,
                          departure_datetime,
                          max_walk_dist,
                          max_trip_duration)

@@

>>>>

- `travel_time_matrix()`: used to compute travel time estimates between one or 
multiple origin/destination pairs for a single departure time or for multiple
departure times over a time window. The output comprises the travel time between
each origin/destination pair. If the user passes a `time_window` parameter
to the function, r5r will calculate multiple travel time matrices departing each 
minute and return additional information on the percentile distribution of travel times
that allows one to grasp service reliability issues.

@@

```{r, message = FALSE, eval = FALSE}
# stable version
install.packages("r5r")
```

@@

@@

To illustrate functionality, the package includes a small sample data set for 
the city of Porto Alegre (Brazil). It includes four files:

- An OSM network in `.pbf` format.
- A public transport network in `GTFS.zip` format.
- The spatial coordinates of points covering the area in `.csv` format.
- The spatial coordinates of some of the city's points of interest.

```{r}
data_path <- system.file("extdata/poa", package = "r5r")
list.files(data_path)

points <- fread(system.file("extdata/poa/poa_hexgrid.csv", package = "r5r"))
points <- points[c(sample(1:nrow(points), 100))]
head(points)
```

@@

Calculating a travel time matrix with r5r can be done very efficiently with the
`travel_time_matrix()` function. The function takes origins/destinations spatial
location (either as a spatial `sf POINT` object, or as a `data.frame` containing
the columns `id`, `lon` and `lat`) and travel parameters as inputs and outputs a
`data.table` with the travel time estimates between all combinations of origins 
and destinations that can be reached with the given parameters.

The `time_window` parameter is set to 1 by default, which means that travel time 
estimates will be calculated using a single departure time (specified by 
`departure_datetime`):

```{r, message = FALSE}
mode <- c("WALK", "TRANSIT")
max_walk_dist <- 1000
max_trip_duration <- 120
departure_datetime <- as.POSIXct("13-03-2019 14:00:00",
                                 format = "%d-%m-%Y %H:%M:%S")

# calculate travel time matrix
ttm <- travel_time_matrix(r5r_core,
                          origins = points,
                          destinations = points,
                          mode = mode,
                          departure_datetime = departure_datetime,
                          max_walk_dist = max_walk_dist,
                          max_trip_duration = max_trip_duration)

@@

Alternatively, a bigger time window can be specified. In such cases several 
travel time matrices are calculated using every minute from `departure_datetime` 
to the end of the given window as a distinct departure time. By default the 
function outputs the 50th percentile (i.e. the median) travel time between each 
origin and destination pair, which can be overruled by the `percentiles` 
parameter:

```{r, message = FALSE}
time_window <- 60
percentiles <- c(25, 50, 75)

# calculate travel time matrix
ttm <- travel_time_matrix(r5r_core,
                          origins = points,
                          destinations = points,
                          mode = mode,
                          departure_datetime = departure_datetime,
                          max_walk_dist = max_walk_dist,
                          max_trip_duration = max_trip_duration,
                          time_window = time_window,
                          percentiles = percentiles)

@@

Since the sample data holds information for few transit routes, many trips rely 
only on walking to be completed. Consequently, the travel time estimates of such 
trips are not affected by transit reliability issues. This is the case of the 
first five rows of `ttm`, which present the same estimate regardless of the 
percentile they represent. The effect of the `time_window` and `percentiles` 
parameters can be observed above in the last rows of `ttm`, in which the 
estimates grow along the increase of the percentile they depict.
